---
title: "IE442 - Fall 2022, Implementation Project"
author: "Ahmet Yiğit Doğan - 2018402105"
date: "14 Jan, 2023"
    
---

# Notes

The below R code runs the SQL calculations to;

- Create a sample database,
- Add the required SQL functions, stored procedures, and triggers to the database,
- And finally insert a sample order to the orders table.

After the SQL calculations, there are some R codes to export final MRP tables after the insertion and the calculations caused by the insertion.

```{r intro, warning = FALSE}

# Required libraries

library("DBI")
library("RMySQL")
library("janitor")  # Will be used for data manipulation after getting results
library("writexl")  # Will be used to export results

# Set working directory to the current folder

setwd(getwd())

# Connecting to database

db <- dbConnect(MySQL(),
                user        = "root",
                password    = "1999",
                host        = "localhost")

# Creating a sample database

dbSendQuery(db, "CREATE DATABASE sample_database;")

# Reconnecting to the database by specifying the database name this time

db <- dbConnect(MySQL(),
                user        = "root",
                password    = "1999",
                dbname      = "sample_database",
                host        = "localhost")

```

# Creating a Sample Database

```{sql connection = db}

CREATE TABLE item ( item_id VARCHAR(16) PRIMARY KEY,
                    item_name VARCHAR(96),
                    lot_size INT NOT NULL,
                    lead_time INT NOT NULL,
                    inventory INT NOT NULL);

```

```{sql connection = db}

INSERT INTO item (item_id, item_name, lot_size, lead_time, inventory)
VALUES  ("001", "Trumpet", 1, 2, 3),
        ("002", "Valve Casing Assembly", 10, 4, 10),
        ("003", "Bell Assembly", 10, 2, 10),
        ("004", "Slide Assemblies", 20, 2, 20),
        ("005", "Valves", 30, 3, 30);

```

```{sql connection = db}

CREATE TABLE bom (  item_id VARCHAR(16),
                    component_id VARCHAR(16),
                    bom_multiplier INT NOT NULL,
                    PRIMARY KEY (item_id, component_id));

```

```{sql connection = db}

INSERT INTO bom (item_id, component_id, bom_multiplier)
VALUES  ("001", "002", 1),
        ("001", "003", 1),
        ("002", "004", 3),
        ("002", "005", 3);  

```

```{sql connection = db}

CREATE TABLE item_period (  item_id VARCHAR(16),
                            period_number INT,
                            gross_requirement INT,
                            projected_inventory INT,
                            planned_order_receipt INT,
                            planned_order_release INT,
                            PRIMARY KEY (item_id, period_number));

```

```{sql connection = db}

INSERT INTO item_period (   item_id, period_number, gross_requirement,
                            projected_inventory, planned_order_receipt,
                            planned_order_release)
VALUES  ("001", 0, 0, 3, 0, 0),         
        ("001", 1, 0, 3, 0, 0),       
        ("001", 2, 0, 3, 0, 0),
        ("001", 3, 0, 3, 0, 0),
        ("001", 4, 0, 3, 0, 0),
        ("001", 5, 0, 3, 0, 0),
        ("001", 6, 0, 3, 0, 0),
        ("001", 7, 0, 3, 0, 0),
        ("001", 8, 0, 3, 0, 0),
        ("001", 9, 0, 3, 0, 0),
        ("001", 10, 0, 3, 0, 0),
        ("001", 11, 0, 3, 0, 0),
        ("001", 12, 0, 3, 0, 0),
        ("002", 0, 0, 10, 0, 0),      
        ("002", 1, 0, 10, 0, 0),
        ("002", 2, 0, 10, 0, 0),
        ("002", 3, 0, 10, 0, 0),
        ("002", 4, 0, 10, 0, 0),
        ("002", 5, 0, 10, 0, 0),
        ("002", 6, 0, 10, 0, 0),
        ("002", 7, 0, 10, 0, 0),
        ("002", 8, 0, 10, 0, 0),
        ("002", 9, 0, 10, 0, 0),
        ("002", 10, 0, 10, 0, 0),
        ("002", 11, 0, 10, 0, 0),
        ("002", 12, 0, 10, 0, 0),
        ("003", 0, 0, 10, 0, 0),        
        ("003", 1, 0, 10, 0, 0),
        ("003", 2, 0, 10, 0, 0),
        ("003", 3, 0, 10, 0, 0),
        ("003", 4, 0, 10, 0, 0),
        ("003", 5, 0, 10, 0, 0),
        ("003", 6, 0, 10, 0, 0),
        ("003", 7, 0, 10, 0, 0),
        ("003", 8, 0, 10, 0, 0),
        ("003", 9, 0, 10, 0, 0),
        ("003", 10, 0, 10, 0, 0),
        ("003", 11, 0, 10, 0, 0),
        ("003", 12, 0, 10, 0, 0),
        ("004", 0, 0, 20, 0, 0),        
        ("004", 1, 0, 20, 0, 0),
        ("004", 2, 0, 20, 0, 0),
        ("004", 3, 0, 20, 0, 0),
        ("004", 4, 0, 20, 0, 0),
        ("004", 5, 0, 20, 0, 0),
        ("004", 6, 0, 20, 0, 0),
        ("004", 7, 0, 20, 0, 0),
        ("004", 8, 0, 20, 0, 0),
        ("004", 9, 0, 20, 0, 0),
        ("004", 10, 0, 20, 0, 0),
        ("004", 11, 0, 20, 0, 0),
        ("004", 12, 0, 20, 0, 0),
        ("005", 0, 0, 30, 0, 0),       
        ("005", 1, 0, 30, 0, 0),
        ("005", 2, 0, 30, 0, 0),
        ("005", 3, 0, 30, 0, 0),
        ("005", 4, 0, 30, 0, 0),
        ("005", 5, 0, 30, 0, 0),
        ("005", 6, 0, 30, 0, 0),
        ("005", 7, 0, 30, 0, 0),
        ("005", 8, 0, 30, 0, 0),
        ("005", 9, 0, 30, 0, 0),
        ("005", 10, 0, 30, 0, 0),
        ("005", 11, 0, 30, 0, 0),
        ("005", 12, 0, 30, 0, 0);


```

```{sql connection = db}

CREATE TABLE orders (   order_id VARCHAR(16) PRIMARY KEY,
                        customer_id VARCHAR(16),
                        period_number INT NOT NULL,
                        item_id VARCHAR(16) NOT NULL,
                        amount INT NOT NULL,
                        order_date DATETIME);

```

```{sql connection = db}

CREATE TABLE periods (  period_number INT PRIMARY KEY,
                        start_date DATE,
                        end_date DATE);

```

```{sql connection = db}

INSERT INTO periods (period_number)
VALUES  (1), (2), (3), (4), (5), (6),
        (7), (8), (9), (10), (11), (12);

```

```{sql connection = db}

CREATE TABLE customer ( customer_id VARCHAR(16) PRIMARY KEY,
                        customer_name VARCHAR(16));

```

```{sql connection = db}

CREATE TABLE required_items (   required_item VARCHAR(16),
                                required_amount INT,
                                due_period INT);

```

# SQL Function Definitions

```{sql connection = db}

SET GLOBAL log_bin_trust_function_creators = 1;

```

## GetLeadTime

```{sql connection = db}

CREATE FUNCTION GetLeadTime(required_item_id VARCHAR(16))
	RETURNS INT

	BEGIN

		DECLARE LT INT;

		SET LT = (	SELECT item.lead_time
					FROM item
					WHERE item.item_id = required_item_id);

		RETURN LT;

	END;

```

## GetLotSize

```{sql connection = db}

CREATE FUNCTION GetLotSize(required_item_id VARCHAR(16))
	RETURNS INT

	BEGIN

		DECLARE LS INT;

		SET LS = (	SELECT item.lot_size
					FROM item
					WHERE item.item_id = required_item_id);

		RETURN LS;

	END;

```

## GetProjectedInventory

```{sql connection = db}

CREATE FUNCTION GetProjectedInventory(  required_item_id VARCHAR(16),
                                        required_item_period INT)
	RETURNS INT

	BEGIN

		DECLARE PInv INT;

		SET PInv = (SELECT item_period.projected_inventory
					FROM item_period
					WHERE   item_period.item_id = required_item_id AND
							item_period.period_number = required_item_period);

		RETURN PInv;

	END;

```

# Stored Procedure Definitions

## GetRequiredItemsCount

```{sql connection = db}

CREATE PROCEDURE GetRequiredItemsCount(
	IN item VARCHAR(16),
	IN parent_item_count INT,
    IN parent_item_due INT)

	BEGIN
        -- CLEAR THE REQUIRED ITEMS TABLE
 		DELETE FROM required_items;
        
        INSERT INTO required_items (required_item, required_amount, due_period)
        
            WITH RECURSIVE cte (it, ad, due) AS (
                SELECT DISTINCT item_id, parent_item_count, parent_item_due
                FROM bom WHERE item_id = item
                UNION ALL
                SELECT bom.component_id,  cte.ad*bom.bom_multiplier, cte.due - GetLeadTime(cte.it) 
                FROM cte JOIN bom ON cte.it = bom.item_id
            )
            SELECT * FROM cte ORDER BY it;

	END;

```

## UpdateMRPTables

```{sql connection = db}

CREATE PROCEDURE UpdateMRPTables ()
    BEGIN
        DECLARE done_i INT DEFAULT 0;

        DECLARE var_i_id VARCHAR(16);
        
        DECLARE cursor_i CURSOR FOR
        SELECT item_id FROM item_period GROUP BY item_id;

        DECLARE CONTINUE HANDLER FOR NOT FOUND SET done_i = 1;

        OPEN cursor_i;
        
            item_loop: LOOP
                FETCH cursor_i INTO var_i_id;
                
                IF done_i = 1 THEN
                    LEAVE item_loop;
                END IF;

                itemwise_update_block: BEGIN

                    DECLARE done_j INT DEFAULT 0;

                    DECLARE var_j_period INT;
                    DECLARE var_j_gross INT;
                    DECLARE var_j_inventory INT;
                    DECLARE var_j_receipt INT;
                    DECLARE var_j_release INT;

                    DECLARE net_requirement INT;
                    DECLARE receipt_increment INT;

                    DECLARE cursor_j CURSOR FOR
                    SELECT  period_number, gross_requirement,
                            projected_inventory, planned_order_receipt,
                            planned_order_release
                    FROM item_period
                    WHERE   item_id = var_i_id AND
                            period_number <> 0;

                    DECLARE CONTINUE HANDLER FOR NOT FOUND SET done_j = 1;

                    OPEN cursor_j; 

                        period_loop: LOOP
                            FETCH cursor_j INTO var_j_period, var_j_gross,
                                                var_j_inventory, var_j_receipt, 
                                                var_j_release;

                            IF done_j = 1 THEN
                                LEAVE period_loop;
                            END IF;
                        
                            SET net_requirement =   var_j_gross - 
                                                    GetProjectedInventory(  var_i_id,
                                                                            var_j_period - 1);

                            
                            IF net_requirement > 0 THEN
                                SET receipt_increment = CEILING(net_requirement/
                                                                GetLotSize(var_i_id))*
                                                                GetLotSize(var_i_id);

                                UPDATE item_period
                                    SET planned_order_receipt = planned_order_receipt +
                                                                receipt_increment,
                                        projected_inventory =   projected_inventory +
                                                                receipt_increment -
                                                                gross_requirement
                                    WHERE   item_id = var_i_id AND
                                            period_number = var_j_period;

                                UPDATE item_period
                                    SET planned_order_release = planned_order_release +
                                                                receipt_increment
                                    WHERE   item_id = var_i_id AND
                                            period_number = var_j_period - GetLeadTime(var_i_id);

                            ELSE
                                UPDATE item_period
                                    SET projected_inventory = -net_requirement
                                    WHERE   item_id = var_i_id AND
                                            period_number = var_j_period;

                            END IF;
 
                        END LOOP period_loop;
                    CLOSE cursor_j;
                END itemwise_update_block;


            END LOOP item_loop;
        CLOSE cursor_i;
    END;

```

# Definition of the Order Insertion Trigger

```{sql connection = db}

CREATE TRIGGER after_order_insertion AFTER INSERT ON orders
    FOR EACH ROW
    BEGIN

        -- UPDATE GROSS REQUIREMENT

        CALL GetRequiredItemsCount(NEW.item_id, NEW.amount, NEW.period_number);

        UPDATE item_period
        INNER JOIN required_items ON    item_period.item_id = required_items.required_item AND
                                        item_period.period_number = required_items.due_period
        SET item_period.gross_requirement = item_period.gross_requirement +
                                            required_items.required_amount;

        -- UPDATE MRP TABLES BY PROCEDURE CALL

        CALL UpdateMRPTables();

    END;

```

# Test Case

Inserting an order of 15 trumpets, due: 10th week

```{sql connection = db}

INSERT INTO orders (order_id, period_number, item_id, amount, order_date)
VALUES ("ORD001", 10, "001", 15, NOW());

```

# Exporting the Resulting MRP Tables

```{r}

# Exporting results as xlsx files

query <- dbSendQuery(db, "SELECT * FROM item_period;")

results_df  <- dbFetch(query)

for (i in c(1:5)) {
    
    temp <- t(results_df[as.numeric(results_df$item_id) == i, -1]) %>%
        row_to_names(row_number = 1)
    
    write_xlsx(as.data.frame(temp), paste("item", i, "mrp.xlsx", sep = ""))
    
}

```


